---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- badges: start -->
[![GitHub issues](https://img.shields.io/github/issues/Minotau-R/MultiFactor)](https://github.com/Minotau-R/MultiFactor/issues)
[![GitHub pulls](https://img.shields.io/github/issues-pr/Minotau-R/MultiFactor)](https://github.com/Minotau-R/MultiFactor/pulls)
[![R BiocCheck](https://github.com/Minotau-R/MultiFactor/actions/workflows/test.yml/badge.svg)](https://github.com/Minotau-R/MultiFactor/actions/workflows/test.yml)
<!-- badges: end -->

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "man/figures/README-",
    out.width = "100%",
    tidy.opts = list(width.cutoff = 80),
    tidy = TRUE
)
```

# MultiFactor: R package for MultiFactor and LinkMap S7 container classes.

## Introduction
`MultiFactor` and `LinkMap` are a pair of S7 classes to organize and manage 
multiple sets of factors, for instance when tracing or converting feature IDs 
across databases. Methods for `MultiFactor` aim to follow `factor` behaviour.

## Installation instructions

Get the latest stable `R` release from [CRAN](http://cran.r-project.org/). Then
install `MultiFactor` from [GitHub](https://github.com/minotau-R/MultiFactor) 
with `remotes`: 

```{r, 'remotes-install', eval = FALSE}
install.packages("remotes")
remotes::install_github("minotau-R/MultiFactor")
```
