---
title: "MultiFactor"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MultiFactor}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r 'knitr-setup', include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview

MultiFactor aims to provide a consistent toolkit to incorporate relational data 
into data analytical tools and methods. In practice, we expect MultiFactor to be
used in combination with additional packages that extend it. For instance, see
[ariadne](https://github.com/minotau-R/ariadne) and
[anansi](https://github.com/thomazbastiaanssen/anansi), two such packages that 
make use of MultiFactor to link and convert across biological database IDs and
perform muti-omics integration analysis, respectively. 

# Get started using MultiFactor

## Installation instructions

Get the latest stable `R` release from [CRAN](http://cran.r-project.org/). Then
install the released version of `MultiFactor`:

```{r 'cran-install', eval = FALSE}
install.packages("MultiFactor")
```

Or install the development version from this repository:

```{r, 'remotes-install', eval = FALSE}
install.packages("remotes")
remotes::install_github("minotau-R/MultiFactor")
```

## Setup 

```{r 'setup-multifactor'}
library(MultiFactor)
```
