---
title: "MultiFactor"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MultiFactor}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r 'knitr-setup', include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview

`MultiFactor` aims to provide a consistent toolkit to incorporate relational 
data into data-analytical tools and methods. In practice, we expect 
`MultiFactor` to be used in combination with additional packages that extend it.
For instance, see [ariadne](https://github.com/minotau-R/ariadne) and
[anansi](https://github.com/thomazbastiaanssen/anansi), two such packages that 
make use of `MultiFactor` to link and convert across biological database IDs and
perform multi-omics integration analysis, respectively. 

# Get started using MultiFactor

## Installation instructions

Get the latest stable `R` release from [CRAN](http://cran.r-project.org/). Then
install the released version of `MultiFactor`:

```{r 'cran-install', eval = FALSE}
install.packages("MultiFactor")
```

Or install the development version from this repository:

```{r, 'remotes-install', eval = FALSE}
install.packages("remotes")
remotes::install_github("minotau-R/MultiFactor")
```

## Setup 

```{r 'setup-multifactor'}
library(MultiFactor)

# Load demo data
set.seed(010292)
tp <- trade_posts()
```

# MultiFactor and LinkMap objects
## LinkMap
The basic object in the `MultiFactor` package is the `LinkMap`, which contains 
relational information across two types of features. This format is sometimes 
referred to as an [edge list](https://en.wikipedia.org/wiki/Edge_list). 
Under a thin layer of S7, the `LinkMap` object is a `data.frame` where the first
two columns indicate which feature IDs that are considered linked. Optional 
additional columns are considered metadata. 

Let's use an example data set to take a closer look at `LinkMap`. The 
`trade_posts()` function generates data about fictional trade posts, where one 
type of good is traded for another. In this particular case, types of furniture 
are traded for types of utensils. 
```{r 'linkmap-object-start'}
linkmap <- tp[[1]]
linkmap
```
The two types of goods are captured by the two columns, with column names 
representing the *types* of good and rows representing which specific 
*type-pairs* are traded at that trade post, or, more generally, are linked in 
that `LinkMap`. 
```{r 'linkmap-object-lvs'}
levels( linkmap )
```
Notice that `LinkMap` is two `factor` columns in a trench coat. We can inspect 
the levels as usual. 

We can see that we have several of these types of trade posts in our data set. 
However, we also notice that the column names - the types of goods - can differ. 
```{r 'linkmap-object-more'}
tp[[2]]
tp[[3]]
```

## MultiFactor
The `MultiFactor` is in essence a collection of `LinkMap`s. Where `LinkMap`s 
contain relational information across two data types, `MultiFactor`s are a 
representation of the relational graph formed by combining those `LinkMaps`. 
The full object `tp`, which contains the `LinkMaps` we just investigated, is in
fact a `MultiFactor` object: 

```{r 'multifactor-object-start'}
tp
```
Notice that a `MultiFactor` summarizes information across the component 
`LinkMaps` in several ways. First, The matrix shows the types of goods as 
columns and the component `LinkMaps` that contain this information as rows. The 
numbers in the matrix then show the number of unique types of that type of good 
in that particular `LinkMap` - and that are therefore linked to the second feature
in the `LinkMap`. 

For instance, in the top-left corner of the matrix, we can see that the first 
`LinkMap` links furniture and utensils and is aptly named `furniture2utensils`.
Notice that five unique types of furniture as well as five types of utensils are
mentioned in `furniture2utensils`.  

Though `MultiFactor` is a list of `LinkMap`s under the hood, we can also 
interact with its matrix representation, as well as access the component levels:  
```{r 'multifactor-object-matrix'}
dim(tp)
dimnames(tp)
levels(tp)
lengths(levels(tp))
```
Note that levels of the same type are automatically unified across all component
LinkMaps with that type of level.

# weave() a path
Going back to our trading post example, let's say we'd want to get some new 
chairs, but we only have fruit. While both fruit and furniture are available in 
our data set, there aren't any trade post that deals in that particular pair of 
goods. 
However, we could imagine trading our fruit for furniture in steps: We can trade
our fruit for a different good, which we trade for another one, until we reach a
good that we can trade for furniture. The `weave()` function does exactly this: 
```{r 'weave'}
fruit2furniture <- weave(tp, fruit ~ furniture)
fruit2furniture
```
We receive a new `LinkMap` containing all fruits that could be traded for 
furniture. 

## Selecting a path
We can use `select_path()` to find the types of traded goods in order,
or more generally, the paths that were traversed. If several paths exist, all 
will be traversed and included into one `LinkMap`. 
```{r 'weave-sp'}
select_path(tp, fruit ~ furniture)
```

# Compatibility with `igraph` and `Matrix`
## Convert `MultiFactor` and `LinkMap` objects to `igraph` representation 
`MultiFactor` relies heavily on the excellent [`igraph`](https://r.igraph.org/) 
and [`Matrix`](https://matrix.r-forge.r-project.org/) packages, particularly for
path finding and sparse matrix representation. 

### Convert `MultiFactor` main graph representation to `igraph` object
```{r 'igraph-multifactor-sp', fig.height=6, fig.width = 6}
library(igraph)
# Convert to an igraph object
g <- as.igraph(tp)
g
# Plot graph across data types
plot(g)
```

### Convert `LinkMap` relational information to `igraph` object
```{r 'igraph-linkmap-sp', fig.height=6, fig.width = 6}
# Convert to an igraph object
lg <- as.igraph(fruit2furniture)
lg
# Same information as the LinkMap: 
plot(lg)
```

# Convert `LinkMap` linkage information to sparse adjacency `Matrix`
```{r 'Matrix-sp'}
# Convert to an igraph object
m <- as.matrix(fruit2furniture, dimnames = levels(fruit2furniture))
m
```

# Utilities
`MultiFactor` also provides some utility functions for data wrangling on tables 
with features (rows) of the appropriate type. 

```{r 'prep-apply'}
# Generate small example table
furniture_table <- as.data.frame(
  replicate( 10, rbinom(6, rbinom(6, 100, runif(6)), runif(6)) ) 
  )
rownames(furniture_table) <- levels(tp)$furniture
furniture_table
```

## subgroup_apply
`subgroup_apply` allows us to run arbitrary code on subsets of a table, based on
groupings on the left hand of the formula:  
```{r 'subgroup-apply'}
subgroup_apply(
  X = furniture_table, LINK = tp, BY = fruit ~ furniture, FUN = as.data.frame
  )

# More complex example: 
# For all subgroups of furniture corresponding to one particular food, if that 
# group has more than two rows (types of furniture), fit a statistical model. 
# 
subgroup_apply(
  furniture_table, tp, fruit ~ furniture, FUN = function(x) {
    if( NROW(x) <= 2) return( NULL )
    # else: 
    summary( lm(V1 ~ V2, data = x) )
    }
)
```

## Compatibility with the tidyverse: `subgroup_to_tbl`
For those who prefer to use `tidyverse`, `subgroup_to_tbl()` returns a tidy 
wide-format table, ready to be grouped based on the first two columns. 
```{r 'subgroup-tbl'}
subgroup_to_tbl(furniture_table, tp, fruit ~ furniture)
```

## Reading and parsing adjaceny list-formatted files
Relational data between two types (bipartite) is sometimes expressed as a text 
file to be read row-wise. The first element of each row specifies a level of the 
first type of data, whereas all following elements all from the second type, 
indicate which levels it is connected to. These files can be tricky to load and 
properly parse into R. The function `read_adjacency_list()` allows these files 
to be read from file (or URL). 

### Example of adjacency list formatted data
We'll format the `LinkMap` from fruit to furniture that we generated above. 
```{r}
# Use igraph to deconstruct a graph into an adjacency list, coerce to characters
adj <- lapply(
  as_adj_list(as.igraph(fruit2furniture)), as_ids
  )
# Only keep the 'fruit' nodes that have more than one link for now
adj <- adj[ levels(fruit2furniture)$fruit ]
adj <- adj[ lengths(adj) > 0 ]
# collapse names to first elements of character vectors
adj <- mapply(c, names(adj), adj, USE.NAMES = FALSE)
adj <- vapply(adj, paste, collapse = "\t", "")
```

If stored in a file in adjacency list format, it may look like this: 
```
pear\tdesk
cherry\tchair\ttable\tdesk
melon\tchair\ttable\tdesk 
blueberry\tchair\tdesk
```

```{r 'read-adj', fig.height=6, fig.width = 6}
# Create some temporary file to read from:
temp <- tempfile()
# fill the temp file with the adjacency list we just constructed 
t.con <- file(temp, "w")
cat(adj, file = t.con, sep = "\n")
close(t.con)

adj_data <- read_adjacency_list(temp)
adj_data

# Notice that the unlinked data types are no longer in the graph.  
plot(as.igraph(adj_data))
```

# Session Info
```{r 'session-info'}
sessionInfo()
```
