---
title: "MultiFactor"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MultiFactor}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r 'knitr-setup', include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview

MultiFactor aims to provide a consistent toolkit to incorporate relational data 
into data analytical tools and methods. In practice, we expect MultiFactor to be
used in combination with additional packages that extend it. For instance, see
[ariadne](https://github.com/minotau-R/ariadne) and
[anansi](https://github.com/thomazbastiaanssen/anansi), two such packages that 
make use of MultiFactor to link and convert across biological database IDs and
perform muti-omics integration analysis, respectively. 

# Get started using MultiFactor

## Installation instructions

Get the latest stable `R` release from [CRAN](http://cran.r-project.org/). Then
install the released version of `MultiFactor`:

```{r 'cran-install', eval = FALSE}
install.packages("MultiFactor")
```

Or install the development version from this repository:

```{r, 'remotes-install', eval = FALSE}
install.packages("remotes")
remotes::install_github("minotau-R/MultiFactor")
```

## Setup 

```{r 'setup-multifactor'}
library(MultiFactor)
```

## MultiFactor and LinkMap objects
### LinkMap
The basic object in the `MultiFactor` package is the `LinkMap`, which contains 
relational information across two types of features. This format is sometimes 
referred to as an [edge list](https://en.wikipedia.org/wiki/Edge_list). 
Under a thin layer of S7, the `LinkMap` object is a `data.frame` where the first
two columns indicate which feature IDs that are considered linked. Optional 
additional columns are considered metadata. 

```{r 'linkmap-object'}

```

## MultiFactor
The `MultiFactor` is in essence a collection of `LinkMap`s. Where `LinkMap`s 
contain relational information across two data types, `MultiFactor`s are a 
representation of the relational graph formed by combining those `LinkMaps`. 

```{r 'multifactor-object'}

```

# weave() a path

```{r 'weave'}

```

# Compatibility with `igraph` and `Matrix`

# Utilities

## subgroup_apply

```{r 'subgroup-apply'}

```

## tidy/subgroup_to_tbl

```{r 'subgroup-tbl'}

```
